Gravitational Lens Modeling in a Citizen Science Context

### 2014-08-13 text/ms.tex
slight changes
corrected text on RE bias

\author[Küng et al]{Rafael Küng,$^{1}$
Prasenjit Saha,$^{1}$
Elisabeth Baeten,$^{2}$
Claude Cornen,$^{2}$
\newauthor
Christine McMillan,$^{2}$
Julianne Wilcox,$^{2}$
Jonas Odermatt,$^{3}$
Jonathan Coles,$^{4}$
\newauthor
Aprajita Verma,$^{5}$
Anupreeta More,$^{6}$
Surhud More,$^{6}$
and Phil Marshall$^{7}$ \\
$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
$^{2}$The Space Warps Collaboration\\
$^{3}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
$^{4}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
$^{5}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
$^{6}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
$^{7}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
}

@@

An implementation of this method, \spl is tested in a modelling
challenge using 29 simulated lenses drawn from the set created
for the \sw citizen-science lens search.  We find that
volunteers from this online community asserted the image
parities and time ordering consistently in some lenses, but made
errors in other lenses depending on the image morphology. While
errors in image parity and time ordering lead to  large errors in the
mass distribution, the enclosed mass was found to be more
robust: the model-derived Einstein radii found by the volunteers were
consistent with those produced by one of the professional team,
suggesting that given the appropriate tools, gravitational lens
modelling is a data analysis activity that can be crowd-sourced to
good effect. Shortcomings of the software implementation 
include (a)~lens profiles that tend to be systematically shallower
than the lenses (leading to systematic over-estimation of the
Einstein radius  by some $\sim25\%$, and (b)~lack of detailed
arc modelling.  Ideas for improvement are discussed.

@@

% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
% THIS PARAGRAPH TO BE REMOVED WHEN THE FIGURES ARE FIXED!
%
% In these figures, the scales and alignments in the six panels are not
% completely consistent: the three panels on the right have the same
% coordinates; the middle-left and lower-left panels are consistent with
% each other, but only approximately aligned with the right column,
% because the latter are affected by errors in mouse placements; the
% upper-left panel is zoomed out slightly, to accommodate the full
% markup.  For qualitative features, which is all that
% Figures~\ref{fig:6941}--\ref{fig:7022} try to convey, the slightly
% varying scales and alignments are harmless.
% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

@@

@@

\begin{figure*}
  \centering
    \includegraphics[width=0.90\linewidth]{fig/eR_5}
  \caption{Model-recovered versus actual Einstein radii \ERf[,rec] and
    \ERf[,act].
%    Circles denote models where the image parities and
%    time-orderings were correct, with filled circles for models by an
%    expert.  Crosses mark models where the time-ordering was
%    incorrect.
    Plus signs indicate models flagged by the modeller as
    failures by commenting negatively about it in the forum.}
  \label{fig:ER_per_sim}
\end{figure*}

% TODO: update figure, fix caption.

\Figref{ER_per_sim} shows that \ERf[,rec] of the models tend to be too
high.  However, this is entirely due to the GLASS model density
profiles being too shallow, as illustrated above. We can separate out
the performance of the \spl interface and its users by comparing their
results with the  Einstein radii of \spl models made by an expert
(PS). Discounting the models which were flagged by the volunteers as
poor, the mean Einstein radius overestimate was 25\%, with a 20\%
standard deviation (shown by the light grey band in \Figref{ER_per_sim}). 
The expert models show a statistically indistinguishable bias of 20\%,
with standard deviation 8\% (the dark grey band in \Figref{ER_per_sim}).
One source of this systematic error is that it is difficult to center
of the lens accurately: an offset leads to a flatter mass profile for
the model compared to the simulation.

@@

The \spl program was developed by Kueng, with design suggestions from
Coles, Cornen and Saha, and feedback from all co-authors.  The
simulations were created by A.~More, in consultation with Marshall,
S.~More and Verma.  Modelling was done by Baeten, Cornen, McMillan,
Odermatt, Saha and Wilcox, with post-modelling analysis by Kueng and
Saha.  All authors participated in writing and editing the manuscript.

>>>>

\author[Küng et al]{Rafael Küng,$^{1}$
Prasenjit Saha,$^{1}$
Elisabeth Baeten,$^{2}$
Claude Cornen,$^{2}$
\newauthor
Christine Macmillan,$^{2}$
Julianne Wilcox,$^{2}$
Jonas Odermatt,$^{3}$
Jonathan Coles,$^{4}$
\newauthor
Aprajita Verma,$^{5}$
Anupreeta More,$^{6}$
Surhud More,$^{6}$
and Phil Marshall$^{7}$ \\
$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
$^{2}$The Space Warps Collaboration\\
$^{3}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
$^{4}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
$^{5}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
$^{6}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
$^{7}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
}

@@

An implementation of this method, \spl is tested in a modelling
challenge using 29 simulated lenses drawn from the set created
for the \sw citizen-science lens search.  We find that
volunteers from this online community asserted the image
parities and time ordering consistently in some lenses, but made
errors in other lenses depending on the image morphology. While
errors in image parity and time ordering lead to  large errors in the
mass distribution, the enclosed mass was found to be more
robust: the model-derived Einstein radii found by the volunteers were
consistent with those produced by one of the professional team,
suggesting that given the appropriate tools, gravitational lens
modelling is a data analysis activity that can be crowd-sourced to
good effect. Shortcomings of the software implementation 
include (a)~lens profiles that tend to be systematically shallower
than the lenses, leading to systematic over-estimation of the
Einstein radius by $\sim10\%$, and (b)~lack of detailed
arc modelling.  Ideas for improvement are discussed.

@@

@@

\begin{figure*}
  \centering
    \includegraphics[width=0.90\linewidth]{fig/eR_5}
  \caption{Model-recovered versus actual Einstein radii \ERf[,rec] and
    \ERf[,act].
    Plus signs indicate models flagged by the modeller as
    failures by commenting negatively about it in the forum.}
  \label{fig:ER_per_sim}
\end{figure*}

@@

\Figref{ER_per_sim} shows that \ERf[,rec] of the models tend to be too
high.  However, this is entirely due to the GLASS model density
profiles being too shallow, as illustrated above. We can separate out
the performance of the \spl interface and its users by comparing their
results with the  Einstein radii of \spl models made by an expert
(PS). Discounting the models which were flagged by the volunteers as
poor, the mean Einstein radius overestimate was 10\%, with a 15\%
standard deviation (shown by the light grey band in \Figref{ER_per_sim}). 
The expert models show a similar bias,
with standard deviation 10\% (the dark grey band in \Figref{ER_per_sim}).
One source of this systematic error is that it is difficult to center
of the lens accurately: an offset leads to a flatter mass profile for
the model compared to the simulation.

@@

The \spl program was developed by Küng, with design suggestions from
Coles, Cornen and Saha, and feedback from all co-authors.  The
simulations were created by A.~More, in consultation with Marshall,
S.~More and Verma.  Modelling was done by Baeten, Cornen, McMillan,
Odermatt, Saha and Wilcox, with post-modelling analysis by Kueng and
Saha.  All authors participated in writing and editing the manuscript.
